A New Metric for the Analysis of Swarms using Potential Fields

### 2018-10-17 paper.tex
Spelling and grammar revised up to end of Section II

\begin{abstract}
  % An important criterion in the assessment of swarm coordination algorithms is
  % the extent to which they cause unnecessary movement, i.e. movement which
  % occurs as a side effect of the algorithm rather than from the fulfilment of
  % the intended function of the swarm. 
  There are many metrics defined for the analysis of swarm coordination algorithms. These metrics are usually based upon the distances between agents, the distance between agents and a fixed point or the resultant vector that \textit{potential field} effects produce. This paper examines a distance-based metric that measures a swarm's overall structure using inter-agent distances and introduces a new metric that identifies a swarm's state based upon the resultant magnitude of the vectors produced by the inter-agent interactions that create the agent distribution within a swarm's structure. The algorithms used to implement the swarming feature are based upon cohesion and repulsion vectors between an agent and its neighbours. In comparing and contrasting the two metrics, we find that the \textit{cohesion/repulsion} metric offers a number of advantages over the distance metric. In particular, the \textit{cohesion/repulsion} metric allows the identification of the essential characteristic of a swarm as `expanding', `stable', or `contracting'. These state cannot be identified using a distance-based metric. Practical swarming applications where the new metric can be applied advantageously include area filling and reconnaissance.
\end{abstract}

@@

%\section{Introduction\label{methods:SwarmStability}}
\textit{Swarming} in the animal kingdom of ants, bees, fish and birds for instance has long been studied by scientists. From these studies mathematical models and algorithms have evolved such as those produced by Reynolds~\cite{REY:87}. The models and algorithms have in turn captured the interest of computing scientists who are interested in applying them to large groups of autonomous mobile \textit{agents} (`robots'). The cooperative coordination of these agents can take many forms such as following a set path~\cite{HCS:09}, existing in a static space~\cite{EP:10, GP:02, GP:04} or foraging as a colony~\cite{HER:11, GK:07}. One of the attributes of swarms that has captured the interest of scientists is that the models and algorithms used to coordinate them are generally sets of simple rules. These simple rules cause the agents to appear to work cooperatively. These rules can also generate movement that does not contribute to the goal of the swarm, this movement is a direct result of the agents attempting to move towards an optimum position within the structure. This internal movement can be identified by analysing the changes in the inter-agent interactions. Measuring this unnecessary movement allows the effectiveness of a swarming algorithm to be evaluated. If the internal movement can be reduced then the stability of the swarm is improved and more of the swarms energy is focused on the goal. Reducing the internal movement also provides a more stable platform for the deployment of sensor arrays. Both the metrics presented in this paper allow this evaluation to be carried out when all the sensor ranges are the same however the distance metric is not able to identify the internal movement accurately when the sensor ranges of each agent vary producing irregular inter-agent distances even though the interactions at the vector level indicate that the agents are in fact stable and optimum. This paper discusses the two metrics and how they differ in their approach to identify these changes. \textit{Distance-based} metrics uses variations in the inter-agent spaces as discussed by many research teams including Navarro et al.~\cite{NIM:09}, Gazi and Passino~\cite{GP:02,GP:04}, Barnes et al.~\cite{BAF:06} and Schneider and Wildermuth~\cite{SW:03}. The new metric uses the resultant magnitudes of the agents' \textit{inter-agent vectors} that are defined by interactions of an agent with its neighbours this allows the ``degree'' of a relationship to be evaluated globally and therefore allows a swarm to be identified as ``expanding'', ``stable'' or ``contracting''.

Many coordination algorithms exist to control the movement of agents within a swarm. The primary goal of these algorithms are to maintain a stable swarm structure. A stable structure is one where the agents in the swarm ``act'' as a single entity. The main focus of this paper is the use of a metric that can be applied to any coordination algorithms that use \textit{potential fields} to produce a swarming effect. Most swarming algorithms define a set of equations that calculate the repulsion and cohesion effects. These two equations are the basis of the new metric.

To facilitate the analysis of the new metric a basic swarm model is introduced and then two metrics are applied to this model to present a comparison of the information provided by a \textit{distance-based} and the new \textit{cohesion/repulsion-based} metric. The analysis includes the identification of a swarm's ``state''. The metric is then applied to a flood filling scenario to highlight a possible use of the new metric.

The remainder of this paper is structured as follows Section~\ref{Section:Related} discusses different metrics that are currently used to analyse swarms, Section~\ref{Section:SwarmModelling} introduces a basic swarm model to produce an environment that can be analysed and evaluated using the two metrics. 
Section~\ref{Section:StabilityMagnitude} discusses the mathematical model that produces the swarming effect and how the potential fields are applied. 
Section~\ref{Section::Movement1} discusses how the model is implemented over time to produce movement. Section~\ref{Section:StabilityNullVector} discusses the effect of the model on swarm stability and the effects models have on swarm movements. 
% Section~\ref{Section:Jitter} discusses the effect the mathematical model can have upon the structure of a swarm in terms of potentially wasted energy. 
Section~\ref{Section:DistanceDynamics} discusses a distance-based metric approach to swarm analysis. Section~\ref{Section:MagnitudeDynamics} introduces the new \textit{cohesion/repulsion-based} metric and how it is calculated and applied. Section~\ref{Section:ApplicationFloodFilling} compares the two metrics using a scenario of a swarm being used to fill an area. Section~\ref{Section:MagnitudeDistanceComparison} discusses the results of the experiments and Section~\ref{Section:FutureWork} discusses additional work that needs to be carried out to exploit the additional information that the new metric provides.

@@

Many metrics exist for the analysis of swarms. These metrics can be categorised as \textit{Centroidal-based}, \textit{Agent/Inter-agent distance-based} and \textit{Magnitude-based}. The common denominator of \textit{centroidal} and \textit{Agent/inter-agent distance} metrics are that they \textit{measure} the agent distribution within a swarm based upon distances. \textit{Magnitude-based} metrics analysis a swarm based upon inter-agent interaction.

@@

Here, ($\bar{r}$) denotes the position vector of the swarm's \textit{centroid}; $r_b$ is the position of each agent($b$) and $S$ is the set of agents that make up the swarm.

Navarro et al. use the centroid to produce a "Mean Distance from the Boundary to the Center of the Group" metric~\cite{NIM:09}. This metric takes the distance of each agent from the centroid and calculates the mean and deviation to produce their measure. Gazi and Passino use this technique when analysing the stability of a swarm based on agent distribution~\cite{GP:11}. The \textit{centroid} is also used as a reference point to determine the path of a swarm. Gazi and Passino use the centroid when analysing the stability of a swarm which is designed to maintain its position by analysing the changes in the centroid's position over time~\cite{GP:04}. Gazi and Passino also use a centroid-based analysis in~\cite{GP:02}~to identify directional changes in agent movement by creating a ``map'' of the potential fields of a swarm in its entirety and overlaying that to the ``centre'' of the swarm.  Barnes et. al. use the centroid to analyse the effects of using bivariate normal functions in their work on using potential fields~\cite{BAF:06, BAFVM:06}.

@@

\textit{Inter-Agent Distance} analysis is a basic mechanism for analysing a swarm in terms of separation of all of the agents with respect to each agent's visible neighbours. This is different to using the \textit{centroidal} approach as the measure is based on agent interactions rather than a resultant structure which effects the centroid. This technique is the most popular approach to analysing a swarms structure. Gazi and Passino use a distance-based analysis in~\cite{GP:04, GP:11} to identify optimum distribution of agents based upon different cohesion and repulsion applications. Wu et. al. discuss using distance-based metrics to evaluate emergent behaviours in swarm based on a Viscsek model~\cite{WCW:11}. These include calculating the average number of neighbours for each agent, grouping agents based on numbers of neighbours (to identify clusters) and the average distance between agents~\cite{WCW:11}. Schneider and Wildermuth~\cite{SW:03} use a slightly different approach to measuring performance. They identify the difference in the paths of robots in a simulated versus a physical implementation and calculate a distance variance between them.

\textit{Magnitude-based} analysis uses the \textit{potential fields} to calculate the ``degree'' of interaction between agents. Bennet and McInnes use the resultant relationship between agents to identify equilibrium states when creating structures such as rings and controlling the direction of agents to created specific transformations~\cite{BM:09}. Wu et al. introduce the concept of a \textit{rejection potential} in an attempt to generate a "kinetic energy" score for a swarm~\cite{WCW:11}.

This paper builds upon the work relating to the \textit{Magnitude-based} metric analysis as discussed by Bennet and McInnes and Wu et al. by expanding the analysis to the entire swarm, deriving a scalar value can be used in the analysis of swarm behaviour.

%% \section{Metric principles\label{section:MagnitudeDynamics}}
%% This section discusses the theory behind swarm movement and the effects that can be measured using both the metrics described above.
\section{Swarm modelling}\label{Section:SwarmModelling}
Currently, much swarm research uses field effects as the method of modelling inter-agent interactions~\cite{BAF:06, BAFVM:06, BM:09, APZDAMC:09, GP:02, GP:04, GP:04a, GP:05, GP:11, MYP:09}. The models usually use two field effects to implement the swarming characteristic. These effects are \textit{cohesion}, to draw agents closer, and \textit{repulsion} to prevent agents colliding. \textit{Potential fields} are ranges around an agent that determine the effect other agents have upon its movement~(Fig.~\ref{methods:FieldEffects}). It is usual for the cohesion field to have a radius $C_b$ which is larger than the repulsion radius $R_b$. When an agent ($b'$) moves into the \textit{cohesion field} of an agent ($b$) then $b'$ is said to be a neighbour of $b$ and is subject to cohesion. When an agent $b'$ moves into the \textit{repulsion field} of $b$ then $b$ is subjected to repulsion. When the repulsion magnitude exceeds the cohesion magnitude the agent will have a tendency to move away from $b'$, i.e. repelled. Obstacles can be introduced to the model, when an agent $b$ moves too close to an obstacle, i.e. within the obstacle repulsion range $O_b$, it has a \textit{repulsion vector} applied to move the agent away from the obstacle.

@@

\section{Inter-agent vector magnitude effect on internal movement}\label{Section:StabilityMagnitude}
Figure~\ref{methods:Stability5} shows the cohesion and repulsion vector contributions to $v_c(b)$ and $v_r(b)$ due to neighbour $b'$, as given in Equation~\ref{eq:FlyToCentre1} and \ref{eq:Repulsion1}. Notice that the vectors are along the line of separation $bb'$. \\

@@

Equation~\ref{eq:FlyToCentre1} identifies the resultant cohesion effect of all the neighbours of $b$. Here, $nbr(b)$ defines the set of all the agents in the swarm that are a neighbour of $b$; a neighbour is any agent that fails within the cohesion field.

@@

Equation~\ref{eq:Repulsion1} identifies the resultant repulsion effect of all the neighbours of $b$ where $rep(.)$ returns a set of all agents that are within the repulsion field.

@@

Using the cohesion and repulsion vectors generated by the relationship of $b$ to its neighbour $b'$ a resultant vector can be calculated. Summing the vectors creates a resultant vector that is the overall affects the agent is being subjected to and provides the direction an agent will move based on the relationship. This is the \textit{inter-agent vector}.

Equation~\ref{eq:InterAgentMovement1} identifies the \textit{inter-agent vector} for agent $b$ with respect to its neighbours.

@@

If a \textit{weighted} model is used then~Equation~\ref{eq:InterAgentMovement1}) becomes~\ref{eq:InterAgentMovement2} where $k_c$ and $k_r$ are weightings to change the effect of each vector as shown in Table~\ref{tab:MetricPhysics1}.

@@

\section{Internal movement and the null vector}\label{Section:StabilityNullVector}
When the two vectors (cohesion and repulsion) have magnitudes that are equal and opposite they produce a null vector, indicating that two agents are optimally spaced for a given set of conditions. Although the agents are at an optimum position when resultants are zero it does not mean the swarm is optimally distributed. If a swarm is in a confined space it is possible for an optimum position to be created where the vector magnitude is affected by a compression effect. This phenomenon is used in the identification of the emergent behaviour of \textit{area flooding}.

@@

% The \textit{`cohesion/repulsion metric'} is therefore the scalar value of the \textit{`agent resultant magnitude'} which is altered based upon the difference between the repulsion and cohesion magnitudes~(\ref{eq:CohesionRepulsion}).
This paper defines the cohesion/repulsion metric scalar for an agent as the length of $v(b)$  as defined in Equation~\ref{eq:InterAgentMovement2}: that is, the weighted effects of cohesion and repulsion by neighbours of the agent. This number carries a negative sign if repulsion dominates and a positive sign if cohesion dominates. This is summarised in Equation~\ref{eq:CohesionRepulsion}.

@@

% \section{Cohesion/repulsion internal movement model}\label{Section:StabilityModel}
Section~\ref{Section:StabilityMagnitude}~discussed calculation of resultant cohesion~(Equation~\ref{eq:FlyToCentre1}) and resultant repulsion~(Equation~\ref{eq:Repulsion1}) vectors for each agent deriving a combined \textit{cohesion/repulsion vector} using as shown in Equation~\ref{eq:InterAgentMovement2}.

@@

The standard deviation associated with this mean is calculated as Equation~\ref{eq:SwarmStabilityQuotientT}.
%%  where $\sigma_p(S)$ is the standard deviation and $\mu_p(S)$ is the mean at the same point in time.

@@

\section{Conclusion}\label{Section:MagnitudeDistanceComparison}
%% The two metrics (distance and \textit{cohesion/repulsion}) both measure the structure of a swarm in terms of the distribution of the agents and deviations from a mean. The \textit{cohesion/repulsion} metric also allows the state of the swarm to be identified.
%% 
%% The distance metric is based upon the physical \textit{distribution} of the agents and the cohesion/repulsion metric is based upon the logical \textit{interaction} of the agents.
Distance-based metrics measure a swarm's structure in terms of the position of agents. This is independent of the algorithm that produces the swarm's structure. It provides an analysis of agent distribution at a point in time but does not allow the identification of the future behaviour of a swarm as `expanding', `stable' or `contracting'.

@@

\section{Future work}\label{Section:FutureWork}
The new metric can be applied to the analysis of the effects of coordination
algorithms upon a swarm's structure. An application to flood filling has
been discussed~(Section~\ref{Section:ApplicationFloodFilling}). The metric can also be applied to identifying the effects of 
`healing' algorithms, e.g. to remove `voids' from a swarm. We plan
to investigate this, and other, practical applications of the metric.

>>>>

\begin{abstract}
  % An important criterion in the assessment of swarm coordination algorithms is
  % the extent to which they cause unnecessary movement, i.e. movement which
  % occurs as a side effect of the algorithm rather than from the fulfilment of
  % the intended function of the swarm. 
  There are many metrics defined for the analysis of swarm coordination
  algorithms. These metrics are usually based upon the distances between
  agents, the distance between agents and a fixed point, or the resultant vectors
  that \textit{potential field} effects produce. This paper examines a
  distance-based metric that measures a swarm's overall structure using
  inter-agent distances. It introduces a new metric that identifies a swarm's
  state based upon the resultant magnitude of the vectors produced by the
  agent interactions that create the agent distribution within the swarm's
  structure. The algorithms used to implement the swarming feature are based
  upon cohesion and repulsion vectors between an agent and its neighbours. In
  comparing and contrasting the two metrics, we find that the
  \textit{cohesion/repulsion} metric offers a number of advantages over the
  distance metric. In particular, the \textit{cohesion/repulsion} metric allows
  the identification of the essential characteristic of a swarm as `expanding',
  `stable', or `contracting'. These states cannot be identified using a
  distance-based metric. Practical swarming applications where the new metric
  can be applied advantageously include area-filling and reconnaissance.
\end{abstract}

@@

%\section{Introduction\label{methods:SwarmStability}}

\textit{Swarming} in the animal kingdom of ants, bees, fish and birds, for
instance, has long been studied by scientists. From these studies, mathematical
models and algorithms have evolved, such as those produced by
Reynolds~\cite{REY:87}. The models and algorithms have in turn captured the
interest of computing scientists, who are interested in applying them to large
groups of autonomous mobile \textit{agents} (`robots'). The cooperative
coordination of these agents can take many forms, such as following a set
path~\cite{HCS:09}, existing in a static space~\cite{EP:10, GP:02, GP:04} or
foraging as a colony~\cite{HER:11, GK:07}. One of the attributes of swarms that
has captured the interest of scientists is that the models and algorithms used
to coordinate them are generally sets of simple rules. These simple rules cause
the agents to appear to work cooperatively. Often, the rules can also generate
movement that does not contribute to the goal of the swarm. This movement is a
direct result of the agents attempting to move towards an optimum position
within the structure. This internal movement can be identified by analysing the
changes in the inter-agent interactions. Measuring this unnecessary movement
allows the effectiveness of a swarming algorithm to be evaluated. If the
internal movement can be reduced, then the stability of the swarm is improved
and more of the swarm's energy is focused on the goal. Reducing the internal
movement also provides a more stable platform for the deployment of sensor
arrays. Both metrics, presented in this paper, allow this evaluation to be
carried out when all the sensor ranges are the same. However, the distance metric
is not able to identify the internal movement accurately when the sensor ranges
of each agent vary, producing irregular inter-agent distances, even though the
interactions at the vector level indicate that the agents are in fact stable
and optimum. This paper discusses the two metrics and how they differ in their
approach to identifying these changes. \textit{Distance-based} metrics use
variations in the inter-agent spaces, as discussed by many research teams
including Navarro et al.~\cite{NIM:09}, Gazi and Passino~\cite{GP:02,GP:04},
Barnes et al.~\cite{BAF:06} and Schneider and Wildermuth~\cite{SW:03}. The new
metric uses the resultant magnitudes of the agents' \textit{inter-agent
vectors} that are defined by interactions of an agent with its neighbours. This
allows the ``degree'' of a relationship to be evaluated globally and therefore
allows a swarm to be identified as ``expanding'', ``stable'' or
``contracting''.

Many coordination algorithms exist to control the movement of agents within a
swarm. The primary goal of these algorithms is to maintain a stable swarm
structure. A stable structure is one where the agents in the swarm ``act'' as a
single entity. The main focus of this paper is on metrics that can be
applied to any coordination algorithms that use \textit{potential fields} to
produce a swarming effect. Most swarming algorithms define a set of equations
that calculate the repulsion and cohesion effects. These equations are the
basis of the new metric.

To facilitate the analysis of the new metric, a basic swarm model is introduced
and then two metrics are applied to this model to present a comparison of the
information provided by a \textit{distance-based} and the new
\textit{cohesion/repulsion-based} metric. The analysis includes the
identification of a swarm's ``state''. The metric is then applied to an 
area-filling scenario to highlight a possible use of the new metric.

The remainder of this paper is structured as follows:
Section~\ref{Section:Related} discusses different metrics that are currently
used to analyse swarms; Section~\ref{Section:SwarmModelling} introduces a basic
swarm model to produce an environment that can be analysed and evaluated using
the two metrics; Section~\ref{Section:StabilityMagnitude} discusses the
mathematical model that produces the swarming effect and how the potential
fields are applied; Section~\ref{Section::Movement1} discusses how the model is
implemented over time to produce movement;
Section~\ref{Section:StabilityNullVector} discusses the effect of the model on
swarm stability and the effects models have on swarm movements; 
% Section~\ref{Section:Jitter} discusses the effect the mathematical model can
% have upon the structure of a swarm in terms of potentially wasted energy. 
Section~\ref{Section:DistanceDynamics} discusses a distance-based metric for
swarm analysis; Section~\ref{Section:MagnitudeDynamics} introduces the new
\textit{cohesion/repulsion-based} metric, showing how it is calculated and
applied; Section~\ref{Section:ApplicationFloodFilling} compares the two metrics
using an area-filling scenario;
Section~\ref{Section:MagnitudeDistanceComparison} discusses the results of
experiments, and, finally, Section~\ref{Section:FutureWork} discusses
additional work that needs to be carried out to exploit the additional
information that the new metric provides.

@@

Many metrics exist for the analysis of swarms. These metrics can be categorised
as \textit{centroidal-based}, \textit{agent/inter-agent distance-based} and
\textit{magnitude-based}. The common denominator of \textit{centroidal} and
\textit{agent/inter-agent distance} metrics is that they \textit{measure} the
agent distribution within a swarm based upon distances.
\textit{Magnitude-based} metrics analyse a swarm based upon inter-agent
interaction.

@@

Here, $\bar{r}$ denotes the position vector of the swarm's \textit{centroid};
$r_b$ is the position of each agent, $b$, and $S$ is the set of agents that make
up the swarm.

Navarro et al. use the centroid to produce a "Mean Distance from the Boundary
to the Center of the Group" metric~\cite{NIM:09}. This metric takes the
distance of each agent from the centroid and calculates the mean and deviation
to produce their measure. Gazi and Passino use this technique when analysing
the stability of a swarm based on agent distribution~\cite{GP:11}. The
\textit{centroid} is also used as a reference point to determine the path of a
swarm. Gazi and Passino use the centroid when analysing the stability of a
swarm which is designed to maintain its position, by analysing the changes in
the centroid's position over time~\cite{GP:04}. Gazi and Passino also use a
centroid-based analysis in~\cite{GP:02}~to identify directional changes in
agent movement by creating a ``map'' of the potential fields of a swarm in its
entirety and overlaying that to the ``centre'' of the swarm.  Barnes et. al.
use the centroid to analyse the effects of using bivariate normal functions in
their work on using potential fields~\cite{BAF:06, BAFVM:06}.

@@

\textit{Inter-agent distance} analysis is a basic mechanism for analysing a
swarm in terms of separation of all of the agents with respect to each agent's
visible neighbours. This is different to using the \textit{centroidal} approach,
as the measure is based on agent interactions rather than a resultant structure
which affects the centroid. This technique is the most popular approach to
analysing a swarm's structure. Gazi and Passino use a distance-based analysis
in~\cite{GP:04, GP:11} to identify optimum distribution of agents based upon
different cohesion and repulsion applications. Wu et. al. discuss using
distance-based metrics to evaluate emergent behaviours in swarms based on a
Viscsek model~\cite{WCW:11}. These include calculating the average number of
neighbours for each agent, grouping agents based on numbers of neighbours (to
identify clusters), and calculating the average distance between agents~\cite{WCW:11}.
Schneider and Wildermuth~\cite{SW:03} use a slightly different approach to
measuring performance. They identify the difference in the paths of robots in a
simulated versus a physical implementation and calculate a distance variance
between them.

\textit{Magnitude-based} analysis uses the \textit{potential fields} to
calculate the ``degree'' of interaction between agents. Bennet and McInnes use
the resultant relationship between agents to identify equilibrium states when
creating structures such as rings and controlling the direction of agents to
create specific transformations~\cite{BM:09}. Wu et al. introduce the concept
of a \textit{rejection potential} in an attempt to generate a "kinetic energy"
score for a swarm~\cite{WCW:11}.

This paper builds upon the work relating to the \textit{magnitude-based} metric
analysis, as discussed by Bennet and McInnes, and Wu et al., by expanding the
analysis to the entire swarm and deriving a scalar value that can be used in the
analysis of swarm behaviour.

%% \section{Metric principles\label{section:MagnitudeDynamics}}
%% This section discusses the theory behind swarm movement and the effects that can be measured using both the metrics described above.
\section{Swarm modelling}\label{Section:SwarmModelling}
Currently, much swarm research uses field effects as the method of modelling
inter-agent interactions~\cite{BAF:06, BAFVM:06, BM:09, APZDAMC:09, GP:02,
GP:04, GP:04a, GP:05, GP:11, MYP:09}. The models usually use two field effects
to implement the swarming characteristic. These effects are \textit{cohesion},
to draw agents closer, and \textit{repulsion} to prevent agents colliding.
\textit{Potential fields} are ranges around an agent that determine the effect
other agents have upon its movement~(Fig.~\ref{methods:FieldEffects}). It is
usual for the cohesion field to have a radius $C_b$ which is larger than the
repulsion radius $R_b$. When an agent $b'$ moves into the \textit{cohesion
field} of an agent $b$, then $b'$ is said to be a neighbour of $b$, and $b$ is
subject to cohesion. When an agent $b'$ moves into the \textit{repulsion field}
of $b$, then $b$ is subject to repulsion. When the repulsion magnitude exceeds
the cohesion magnitude, the agent $b$ will have a tendency to move away from
$b'$, i.e. to be repelled. Obstacles can be introduced into the model. When an
agent $b$ moves too close to an obstacle, i.e. within the obstacle repulsion
range $O_b$, it has a \textit{repulsion vector} applied, giving it a tendency
to move away from the obstacle.

@@

\section{Inter-agent vector magnitude effect on internal movement}\label{Section:StabilityMagnitude}
Figure~\ref{methods:Stability5} shows the cohesion and repulsion vector
contributions of neighbour $b'$ to $v_c(b)$ and $v_r(b)$, as given in
Equations~\ref{eq:FlyToCentre1} and \ref{eq:Repulsion1}. Notice that the vectors
are along the line of separation $bb'$. \\

@@

Equation~\ref{eq:FlyToCentre1} identifies the resultant cohesion effect of all
the neighbours of $b$. Here, $nbr(b)$ defines the set of all the agents in the
swarm that are a neighbour of $b$; a neighbour is any agent that falls within
the cohesion field.

@@

Equation~\ref{eq:Repulsion1} identifies the resultant repulsion effect of all
the neighbours of $b$, where $rep(b)$ returns the set of all agents that are
within the repulsion field of $b$.

@@

Using the cohesion and repulsion vectors generated by the relationship of $b$
to its neighbours, a resultant vector can be calculated that summarises the
effects to which the agent is being subjected, and provides the direction in
which the agent will move, based on this relationship. This is the
\textit{inter-agent vector}.  Equation~\ref{eq:InterAgentMovement1} identifies
the \textit{inter-agent vector} for agent $b$ with respect to its neighbours.

@@

If a \textit{weighted} model is used then~Equation~\ref{eq:InterAgentMovement1}
is replaced by~Equation~\ref{eq:InterAgentMovement2}, where $k_c$ and $k_r$ are
weightings to change the effect of each vector, as exemplified in
Table~\ref{tab:MetricPhysics1}.

@@

\section{Internal movement and the null vector}\label{Section:StabilityNullVector}

When the two vectors (cohesion and repulsion) have magnitudes that are equal
and opposite they produce a null vector, indicating that two agents are
optimally spaced for a given set of conditions. Although the agents are at an
optimum position when resultants are zero it does not mean the swarm is
optimally distributed. If a swarm is in a confined space it is possible for an
optimum position to be created where the vector magnitude is affected by a
compression effect. This phenomenon is used in the identification of the
emergent behaviour of \textit{area filling}.

@@

% The \textit{`cohesion/repulsion metric'} is therefore the scalar value of the \textit{`agent resultant magnitude'} which is altered based upon the difference between the repulsion and cohesion magnitudes~(\ref{eq:CohesionRepulsion}).

This paper defines the cohesion/repulsion metric scalar for an agent as the
length of $v(b)$  as defined in Equation~\ref{eq:InterAgentMovement2}: that is,
the weighted effects of cohesion and repulsion by neighbours of the agent. This
number carries a negative sign if repulsion dominates and a positive sign if
cohesion dominates. This is summarised in Equation~\ref{eq:CohesionRepulsion}.

@@

% \section{Cohesion/repulsion internal movement model}\label{Section:StabilityModel}

Section~\ref{Section:StabilityMagnitude}~discussed calculation of resultant
cohesion~(Equation~\ref{eq:FlyToCentre1}) and resultant
repulsion~(Equation~\ref{eq:Repulsion1}) vectors for each agent deriving a
combined \textit{cohesion/repulsion vector} using as shown in
Equation~\ref{eq:InterAgentMovement2}.

@@

The standard deviation associated with this mean is calculated as
Equation~\ref{eq:SwarmStabilityQuotientT}.

%%  where $\sigma_p(S)$ is the standard deviation and $\mu_p(S)$ is the mean at the same point in time.

@@

\section{Conclusion}\label{Section:MagnitudeDistanceComparison}
%% The two metrics (distance and \textit{cohesion/repulsion}) both measure the structure of a swarm in terms of the distribution of the agents and deviations from a mean. The \textit{cohesion/repulsion} metric also allows the state of the swarm to be identified.
%% 
%% The distance metric is based upon the physical \textit{distribution} of the agents and the cohesion/repulsion metric is based upon the logical \textit{interaction} of the agents.

Distance-based metrics measure a swarm's structure in terms of the position of
agents. This is independent of the algorithm that produces the swarm's
structure. It provides an analysis of agent distribution at a point in time but
does not allow the identification of the future behaviour of a swarm as
`expanding', `stable' or `contracting'.

@@

\section{Future work}\label{Section:FutureWork}

The new metric can be applied to the analysis of the effects of coordination
algorithms upon a swarm's structure. An application to area filling has been
discussed~(Section~\ref{Section:ApplicationFloodFilling}). The metric can also
be applied to identifying the effects of `healing' algorithms, e.g. to remove
`voids' from a swarm. We plan to investigate this, and other, practical
applications of the metric.
